\documentclass[10pt, twocolumn]{article}
\usepackage[margin=0.75in, top=0.6in, bottom=0.6in]{geometry}
\usepackage{newtxtext}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{xurl}
\usepackage{hyperref}
\usepackage{xcolor}

% Spacing
\setlist{nosep, leftmargin=*} 
\linespread{1.05} 
\setlength{\parskip}{3pt} 
\hyphenpenalty=10000
\exhyphenpenalty=10000
\renewcommand{\footnotesize}{\scriptsize}

% Section formatting with visual distinction
\titleformat{\section}{\normalsize\bfseries\scshape}{\thesection}{0.5em}{}[\vspace{1pt}\titlerule]
\titlespacing*{\section}{0pt}{6pt}{3pt}
\titleformat{\subsection}{\small\bfseries\itshape}{\thesubsection}{0.5em}{}
\titlespacing*{\subsection}{0pt}{4pt}{2pt}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    citecolor=black,
    breaklinks=true,
}

\pagestyle{plain} 

% Compact Bibliography
\makeatletter
\renewenvironment{thebibliography}[1]
     {\section*{References}%
      \scriptsize
      \setlength{\topsep}{0pt}
      \list{\@biblabel{\@arabic\c@enumiv}}%
           {\settowidth\labelwidth{\@biblabel{#1}}%
            \leftmargin\labelwidth
            \advance\leftmargin\labelsep
            \itemsep1pt\parsep0pt\parskip0pt
            \usecounter{enumiv}%
            \let\p@enumiv\@empty
            \renewcommand\theenumiv{\@arabic\c@enumiv}}%
      \sloppy
      \clubpenalty4000
      \@clubpenalty \clubpenalty
      \widowpenalty4000%
      \sfcode`\.\@m}
     {\def\@noitemerr
       {\@latex@warning{Empty `thebibliography' environment}}%
      \endlist}
\makeatother

\begin{document}

\twocolumn[{%
\noindent
\textbf{\Large Memo \#1: Infrastructure for the AI Era} \hfill \textbf{Daniel Hardesty Lewis} \\
\noindent
\textit{Energy Infrastructure: Grid Resilience and Data Center Power Demand} \hfill March 3, 2026 \\
\textbf{PLAN A6613: AI and the Future of Cities} \hfill Spring 2026
\vspace{3pt}
\hrule
\vspace{8pt}
}]

\section*{I. Infrastructure Selection \& The Physical Layer}

The infrastructure type examined in this memo is \textbf{energy}, specifically the electrical grid capacity and resilience required to power AI-driven computation at urban scale. The central argument is that AI energy infrastructure creates a \textit{temporal lock-in}, a concept rooted in Unruh's (2000) analysis of how large-scale infrastructure investments in energy systems generate self-reinforcing dependencies that resist change: rapid GPU obsolescence cycles drive urgent demand for power capacity, locking cities into long-term grid commitments and fiscal dependencies that progressively reduce municipal leverage over the infrastructure's most consequential impacts. The tangible hardware includes high-voltage transmission lines, distribution substations, transformers, backup generators, battery energy storage systems, and the power plants themselves. Within data centers, the physical layer extends to uninterruptible power supplies, power distribution units, and cooling systems that consume a significant share of total facility energy. Cooling accounts for roughly 7\% of electricity at efficient hyperscale facilities but exceeds 30\% at less efficient enterprise data centers, and mid-sized facilities can consume up to 300,000 gallons of water daily for this purpose alone (Leppert, 2025; NLC, 2024).

These assets are physically located along two axes. \textit{Generation} occurs at power plants, increasingly supplemented by on-site solar arrays and, prospectively, small modular nuclear reactors, often sited tens or hundreds of miles from the cities they serve. Several tech companies have announced purchasing agreements with nuclear power startups, and plans are underway to revive retired nuclear plants like Three Mile Island in Pennsylvania specifically to meet data center demand (Leppert, 2025). \textit{Distribution} infrastructure threads through urban rights-of-way, with substations occupying industrial parcels adjacent to data center campuses. Northern Virginia's Data Center Alley along Route 28 in Loudoun County concentrates over 300 data centers within a single corridor, creating localized demand that exceeds the capacity originally engineered for a suburban residential load profile (Northern Virginia Magazine, 2024). Nearly half of all U.S.\ data center capacity sits in just five regional clusters, and 50\% of facilities under development are being built within these same pre-existing concentrations, raising risks of local grid bottlenecks (IEA, 2025).

This infrastructure is necessary because every AI workload ultimately resolves to electricity consumed by GPUs in a rack. A typical AI-focused data center consumes as much electricity as 100,000 households; the largest facilities under construction today will consume twenty times that amount (IEA, 2025). The International Energy Agency estimates that global data center electricity consumption reached 415 terawatt-hours in 2024, approximately 1.5\% of total global electricity consumption, and projects this figure will more than double to 945 TWh by 2030, slightly more than Japan consumes today (IEA, 2025). In the United States alone, data centers consumed 183 TWh in 2024, over 4\% of national electricity use, and accounted for roughly 26\% of Virginia's total electricity supply in 2023 (Leppert, 2025). Data centers now account for nearly half of projected U.S.\ electricity demand growth between now and 2030, meaning the country will soon consume more electricity for data centers than for the production of aluminium, steel, cement, chemicals, and all other energy-intensive goods combined (IEA, 2025). Without sufficient grid capacity, AI-dependent city services cannot function reliably. The scale of this infrastructure challenge raises a governance question: who decides where these facilities are built, who pays for the grid upgrades they require, and who bears the environmental costs?

\section*{II. The Stakeholder Matrix: Objectives and Tensions}

This temporal lock-in becomes concrete when examined through each stakeholder's experience of the same infrastructure.

\subsection*{Local Government}

The first stage of lock-in is fiscal. Cities face a fundamental tension between \textbf{AI for Optimization} and \textbf{AI for Equity}: data center tax revenue arrives immediately, while energy costs accumulate over years. Loudoun County, Virginia, anticipates nearly \$900 million in data center tax revenue in fiscal year 2025, a sum approaching the county's entire operating budget (Lincoln Institute, 2025). In Mecklenburg County, data center taxes account for more than a third of the county's general revenue and helped pay for a \$154 million school renovation project (Cardinal News, 2026). Once a county funds its school system through data center revenue, rejecting future facilities means cutting public services. Yet the same concentration of energy demand threatens to raise electricity prices for residents who derive no direct benefit from AI services. A Carnegie Mellon University study estimates that data center growth could increase average U.S.\ electricity bills by 8\% by 2030, with increases exceeding 25\% in heavily concentrated markets like Northern Virginia (Leppert, 2025).

Denver illustrates one attempt to interrupt this ratchet: in 2024, the city considered a moratorium on new data center construction to give policymakers time to update zoning and environmental review processes before the next round of commitments (Kiowa County Press, 2024). The National League of Cities has called for advance planning requirements before grid expansions are approved, aiming to insert a review stage between a developer's proposal and the irreversible utility commitments that follow (NLC, 2024). Several states have weighed bills requiring data centers to report electricity and water usage before receiving permits, an effort to create the informational preconditions for intervention at the commitment point rather than after (Leppert, 2025).

\subsection*{Businesses}

Data center operators and cloud providers (Amazon Web Services, Microsoft Azure, Google Cloud) make money by selling compute capacity on a consumption or subscription basis. Their business model requires massive upfront capital expenditure on land, power interconnection, and server hardware, amortized over 15--20 year facility lifespans. This creates a structural misalignment with the \textbf{Speed vs.\ Incremental Evolution} tension in AI technology: a GPU generation is obsolete within 2--3 years, but the electrical and cooling infrastructure built to serve it must last decades. Operators manage this tension through modular facility designs that allow incremental GPU upgrades without rebuilding the power envelope (Kiran \& Wu, 2025). Virginia's new GS-5 rate class, which requires 14-year contracts with minimum demand commitments of 85\% of contracted capacity, further locks operators into long-term grid commitments that must outlast multiple hardware generations (American Action Forum, 2025).

A second tension exists between operators' renewable energy commitments and the speed of deployment. Microsoft, Google, and Amazon have each pledged carbon neutrality or 100\% renewable matching, yet natural gas supplied over 40\% of U.S.\ data center electricity in 2024, with renewables accounting for only about 24\%, nuclear about 20\%, and coal about 15\% (Leppert, 2025; IEA, 2025). This gap means that long-term grid commitments made today lock cities into a fossil-fuel-heavy energy mix for decades, regardless of corporate sustainability pledges. The federal government has identified data center development as a national priority and issued executive orders to accelerate permitting, compressing the already-narrow window in which cities might impose conditions before construction begins (Leppert, 2025). Where communities have mobilized early enough, pushback has delayed or blocked billions of dollars in projects, but these successes illustrate a reactive pattern: opposition arises only after commitments have advanced far enough to become visible, by which point the most consequential decisions are already locked in (Kiran \& Wu, 2025).

\subsection*{Residents}

Residents experience the delayed consequences of commitments made before they could weigh in. As \textit{ratepayers}, they absorb electricity price increases driven by utility investments in transmission and distribution upgrades to serve data centers. In the PJM electricity market stretching from Illinois to North Carolina, data centers accounted for an estimated \$9.3 billion price increase in the 2025--26 capacity market, with average residential bills expected to rise by \$18 a month in western Maryland and \$16 a month in Ohio (Leppert, 2025). The Virginia State Corporation Commission approved Dominion Energy rate increases of approximately \$16 per month for typical residential customers beginning in 2026, a 9\% increase over two years (Inside Climate News, 2025). Northern Virginia residents have reported dramatic bill spikes during 2024 as grid infrastructure costs accelerated (ARLnow, 2024). Nationally, the typical U.S.\ household was billed \$142 per month for electricity in 2024, up 25\% from \$114 in 2014, with data center demand compounding already-rising infrastructure costs from extreme weather hardening and cybersecurity upgrades (Leppert, 2025).

As \textit{neighbors}, residents discover the physical costs of data centers only after siting decisions are irreversible: noise from industrial cooling fans, light pollution from 24/7 operations, and construction truck traffic are consequences that materialize once facilities are built but were not salient during the approval process. The Coalition to Protect Prince William County and the Virginia Data Center Reform Coalition formed after construction was already underway or approved, illustrating the temporal gap between commitment and consequence (Northern Virginia Magazine, 2024). Water consumption follows the same pattern: U.S.\ data centers consumed about 17 billion gallons in 2023, with hyperscale facilities projected to consume 16--33 billion gallons annually by 2028 (Leppert, 2025). In Botetourt County, a planned Google data center could draw two to eight million gallons per day (Cardinal News, 2026), locking the region into long-term water allocation commitments that parallel the electrical grid's lock-in.

As \textit{environmental stakeholders}, residents bear the grid reliability risks that accompany concentrated demand. A Department of Energy analysis found that the Northern Virginia region could face over 430 hours of power outages annually by 2030, compared to 2.4 hours currently, if grid capacity fails to keep pace with data center load growth (Northern Virginia Magazine, 2024). These risks determine whether a family's air conditioning works during a heat wave or whether a hospital's backup systems are sufficient during a grid stress event. Residents are not merely end-users of AI services; they are involuntary subsidizers of infrastructure commitments that were made years before these costs materialized. By the time bill increases, water competition, and grid fragility become visible, the 14-year contracts and fiscal dependencies are already locked in.

\section*{III. The Digital Layer \& Governance}

The stakeholder analysis reveals how the temporal lock-in progresses: operators commit to 14-year contracts before cities can assess grid impacts; tax revenue creates fiscal dependency before environmental costs materialize; and the data needed to evaluate these trade-offs remains proprietary throughout. This section examines the data and regulatory structures that prevent cities from intervening before each commitment becomes irreversible.

\subsection*{Data Ownership}

The digital layer of energy infrastructure encompasses grid telemetry data (real-time load, voltage, frequency), building energy management system data, and the operational metrics that data center operators monitor continuously. Ownership is fragmented. Utilities like Dominion Energy own grid-level data and share limited aggregates with regulators; data center operators treat facility-level consumption, cooling efficiency measured as Power Usage Effectiveness, and workload distribution as proprietary trade secrets. Residents who generate solar power or participate in demand-response programs produce data that utilities collect but rarely make accessible in usable form. This is largely a proprietary silo model: no single entity has a comprehensive view of how much energy AI workloads actually consume at the urban scale, which makes evidence-based planning nearly impossible. The IEA itself notes that policy makers and markets have lacked the tools to assess implications of AI's deepening connection to the energy sector, in part because of this data fragmentation (IEA, 2025).

\subsection*{Regulatory Powers}

Energy infrastructure governance is jurisdictionally fragmented in ways that systematically disadvantage cities. \textit{State public utility commissions}, not city governments, set electricity rates, approve new generation capacity, and regulate transmission siting. Virginia's State Corporation Commission, for instance, approved the new GS-5 rate class for large electricity consumers like data centers in 2025, requiring 14-year contracts and minimum demand commitments of 85\% of contracted distribution and transmission capacity (American Action Forum, 2025). \textit{Federal agencies} such as FERC and the Department of Energy govern interstate transmission and wholesale electricity markets, and the White House has issued executive orders to accelerate federal permitting of data center infrastructure (Leppert, 2025). Cities are left with \textit{local zoning and land use} authority: they can determine where data centers are built, impose setback and noise requirements, and negotiate community benefit agreements, but they cannot directly control the price of electricity or mandate renewable sourcing. An Iowa county recently adopted extensive zoning rules specifically in response to its third data center, illustrating how local governments use the limited tools available to them (Inside Climate News, 2026). This jurisdictional structure accelerates the lock-in: by the time city residents experience the costs, the binding decisions have already been made by state commissions and federal agencies operating on timescales and at jurisdictional levels that preclude meaningful municipal input.

\subsection*{Looking Ahead}

If the core problem is temporal, the interventions must target timing. Three mechanisms could interrupt the ratchet. First, \textbf{staged permitting with mandatory review intervals} would replace one-time zoning approval with conditional permits requiring reassessment every five years, roughly aligned with GPU generation cycles. This creates structured breakpoints where cities can renegotiate terms, impose new conditions, or deny expansion before the next round of grid commitments locks in. A Brookings Institution symposium found that communities face ``compressed timelines and asymmetric negotiations'' that leave local leaders unable to ``weigh short-term fiscal gains against long-term infrastructure and public service obligations'' (Pipa \& Aley, 2026); staged review directly decompresses these timelines. Second, \textbf{revenue diversification requirements} would cap data center tax revenue at a fixed share of a county's general fund, requiring localities to maintain fiscal independence before approving additional facilities. Loudoun County's Revenue Stabilization Fund, created in 2023, acknowledges this dependency but does not limit it. A statutory cap would prevent the fiscal lock-in that makes future regulation politically impossible. Third, \textbf{temporal impact assessments} would require cities to project infrastructure costs across all four timescales (technology cycle, contract duration, facility lifespan, and cumulative fiscal dependency) before approving new data center development, using data aggregated through municipal energy data trusts that address the current proprietary information gap. Cost allocation reforms such as Virginia's SB 253, which would shift substation costs from residential ratepayers to data center operators and save households an estimated \$65 annually (13NewsNow, 2025), would complement these mechanisms by reducing the cost burden that residents absorb during the delay between commitment and consequence.\footnote{Full process documentation, including source downloads, annotations, synthesis matrix, verification audit, and prompt log, is available in the project repository at \texttt{memo1/}.}

\begin{thebibliography}{15}
\bibitem{aaf} American Action Forum. (2025, January). Virginia's new data center rate class: Policy brief. \url{https://www.americanactionforum.org/}
\bibitem{aipb} Kiran, M. M. \& Wu, C. (2025, September 17). U.S. states and cities are shaping the future of AI infrastructure. \textit{AI Policy Bulletin}. \url{https://aipolicybulletin.org/articles/us-states-and-cities-are-shaping-the-future-of-ai-infrastructure}
\bibitem{arlnow} ARLnow. (2024). Northern Virginia residents report skyrocketing electricity bills amid data center boom. \textit{ARLnow}. \url{https://www.arlnow.com/}
\bibitem{cardinal2} Cardinal News. (2026, February 25). Data center complex with gas power plant planned for Wise County. \textit{Cardinal News}. \url{https://cardinalnews.org/2026/02/25/data-center-complex-with-gas-power-plant-planned-for-wise-county/}
\bibitem{icn1} Inside Climate News. (2025, November). SCC approves Dominion Energy rate increase for 2026. \textit{Inside Climate News}. \url{https://insideclimatenews.org/}
\bibitem{icn2} Inside Climate News. (2026, March). Facing its third data center, an Iowa county rolls out extensive zoning rules. \textit{Inside Climate News}. \url{https://insideclimatenews.org/}
\bibitem{iea} International Energy Agency. (2025). \textit{Energy and AI}. IEA, Paris. \url{https://www.iea.org/reports/energy-and-ai}
\bibitem{kcp} Kiowa County Press. (2024). Denver considers moratorium on new data centers. \textit{Kiowa County Press}. \url{https://kiowacountypress.net/}
\bibitem{pew} Leppert, R. (2025, October 24). What we know about energy use at U.S. data centers amid the AI boom. \textit{Pew Research Center}. \url{https://www.pewresearch.org/short-reads/2025/10/24/what-we-know-about-energy-use-at-us-data-centers-amid-the-ai-boom/}
\bibitem{news13} 13NewsNow. (2025). Virginia bill would shift data center electrical costs from residential ratepayers. \textit{13NewsNow}. \url{https://www.13newsnow.com/}
\bibitem{nlc} National League of Cities. (2024). Data centers and local environmental considerations. \textit{NLC Research}. \url{https://www.nlc.org/}
\bibitem{novamag} Northern Virginia Magazine. (2024). Data centers: The flashpoint over power demand and infrastructure strain. \textit{Northern Virginia Magazine}. \url{https://northernvirginiamag.com/}
\bibitem{brookings} Pipa, A. F. \& Aley, A. (2026, March 2). The local implications of data centers for rural communities in the US. \textit{Brookings Institution}. \url{https://www.brookings.edu/articles/local-implications-data-centers-rural-communities-us/}
\bibitem{lincoln} Lincoln Institute of Land Policy. (2025). Data centers and local government. \textit{Lincoln Institute}. \url{https://www.lincolninst.edu/}
\bibitem{unruh} Unruh, G. C. (2000). Understanding carbon lock-in. \textit{Energy Policy}, 28(12), 817--830.
\end{thebibliography}

\end{document}
