\documentclass[11pt, twocolumn]{article}
\usepackage[margin=0.65in, columnsep=0.25in, bottom=0.45in]{geometry}
\usepackage{newtxtext}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{xurl}
\usepackage{hyperref}
\usepackage{xcolor}

% Tight compacting
\setlist{nosep, leftmargin=*} 
\linespread{0.93} 
\setlength{\parskip}{2.5pt} 
\renewcommand{\footnotesize}{\tiny}

% Tight section formatting with visual distinction
\titleformat{\section}{\large\bfseries\scshape}{\thesection}{1em}{}[\vspace{1pt}\titlerule]
\titlespacing*{\section}{0pt}{6pt}{3pt}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    citecolor=black,
    breaklinks=true,
}

\pagestyle{plain} 

% Compact Bibliography
\makeatletter
\renewenvironment{thebibliography}[1]
     {\section*{References}%
      \tiny
      \setlength{\topsep}{0pt}
      \list{\@biblabel{\@arabic\c@enumiv}}%
           {\settowidth\labelwidth{\@biblabel{#1}}%
            \leftmargin\labelwidth
            \advance\leftmargin\labelsep
            \itemsep0pt\parsep0pt\parskip0pt
            \usecounter{enumiv}%
            \let\p@enumiv\@empty
            \renewcommand\theenumiv{\@arabic\c@enumiv}}%
      \sloppy
      \clubpenalty4000
      \@clubpenalty \clubpenalty
      \widowpenalty4000%
      \sfcode`\.\@m}
     {\def\@noitemerr
       {\@latex@warning{Empty `thebibliography' environment}}%
      \endlist}
\makeatother

\begin{document}

\twocolumn[
  \begin{@twocolumnfalse}
    \noindent
    \textbf{\LARGE AI Reading Exercise: Class 10} \hfill \textbf{Daniel Hardesty Lewis} \\
    \noindent
    \textit{AI and City Building} \hfill March 31, 2026 \\
    \textbf{PLAN A6613: AI and the Future of Cities} \hfill Spring 2026
    \vspace{3pt}
    \hrule
    \vspace{6pt}
  \end{@twocolumnfalse}
]

\section*{Tool \& Process}
I used \textbf{Google Gemini} with extended thinking via the Antigravity IDE.\footnote{Prompt log, sources, annotations, synthesis matrix, and verification audit archived at \url{https://github.com/dhardestylewis/plan-a6613-ai-reading-class-3/tree/main/week10}.} I stuck with the \textit{source-first} methodology from Weeks 7 and 9, annotating three distinct perspectives (a policy think-tank, a gov-tech vendor, and an industry vendor) before attempting any synthesis. Without this guardrail, the AI essentially wrote an optimistic brochure, regurgitating venture-backed marketing and entirely missing the core structural threat: these tools accelerate capital extraction while aggressively dismantling professional liability and public oversight.

\section*{Key Findings}
\noindent AI is moving into three critical pipelines: site design, permitting, and construction monitoring. By automating the subjective components of professional judgment, the market is aggressively outpacing the democratic regulations meant to constrain it.

\textbf{\textit{1. Site design weaponizes feasibility to enforce algorithmic redlining.}}
TestFit maps site plans based on parameters like developer yield and parking ratios (TestFit, 2024). It runs thousands of layouts in seconds to find the highest and best use. While it integrates with Revit and AutoCAD, the algorithm operates as an impenetrable black box aggressively prioritizing capital extraction. This is the hyper-commodification of space: it threatens to obliterate equitable urban planning in favor of cookie-cutter density, laundering historical spatial inequality through neutral-sounding mathematics (algorithmic redlining). Walker notes AI shouldn't ``replace human professionals,'' but the ease of instant feasibility studies creates an insurmountable leverage asymmetry against outmatched, underfunded public agencies (Walker, 2024).

\textbf{\textit{2. Computational law mandates the neoliberal enclosure of civic governance.}}
Symbium translates zoning codes into logic programming via Complaw\textsuperscript{\textregistered}, bypassing manual review for instant permits (Symbium, 2024). Driven by California's SB 379 mandates for residential solar, Symbium is deployed in San Mateo and Bakersfield. But when complex public laws are converted by a private vendor, the algorithmic interpretation becomes a corporate trade secret. Planners lose the structural ability to audit compliance, quietly conceding total regulatory capture to PropTech firms. Walker warns this tech remains in a disaster-prone hype cycle (Walker, 2024). When permits issue instantly, public oversight is not augmented; it is actively eradicated to serve private efficiency.

\textbf{\textit{3. Computer vision monitors progress via digital Taylorian surveillance.}}
Doxel outfits workers with 360-degree cameras to verify construction progress against BIM models across fundamental trades (Doxel, 2024). Claiming to replace ``subjective, manual estimation and reporting with data-backed, visual evidence,'' it targets high-stakes health care and data centers. Yet this is the frontier of digital Taylorism: labor advocates warn this technology functions as panoptic bossware, expanding worker surveillance and automating punitive disciplinary metrics without union negotiation. Furthermore, Doxel explicitly provides \textit{evidence}, not \textit{decisions}. It flags deviations but refuses to rule on code violations, legally insulating the AI while forcing a downstream human inspector to absorb the actual liability.

\section*{Verification and Reflection}
The AI nailed the basic retrieval. It successfully identified the software, founders, and deployments without hallucinating. But left unconstrained, it blindly treated marketing copy as objective reality. I had to force the model to cross-reference those claims against Walker's adversarial framework to recover the critical planning context. The takeaway is stark: PropTech's relentless drive for frictionless efficiency is a Trojan horse for dismantling civic governance. The tech sector is aggressively automating legally binding, intensely political tasks (feasibility, code compliance, worker subjugation) behind the firewall of proprietary algorithms. By deploying these systems faster than municipalities can mount any democratic defense, the industry is not just replicating the NYC MyCity chatbot's accountability gap: it is actively weaponizing AI to enforce algorithmic redlining, finalize regulatory capture, and permanently erase local democracy.

\begin{thebibliography}{9}
\bibitem{doxel2024} Doxel. (2024). About Doxel: AI progress tracking for construction. \url{https://doxel.ai/about}
\bibitem{symbium2024} Symbium. (2024). Symbium and Complaw\textsuperscript{\textregistered}: Property and Permit Information Portal. Stanford CodeX. \url{https://symbium.com}
\bibitem{testfit2024} TestFit. (2024). About TestFit: Our mission, vision, and team. \url{https://www.testfit.io/about}
\bibitem{walker2024} Walker, R. (2024). City Tech: 20 Apps, Ideas, and Innovators Changing the Urban Landscape. Lincoln Institute of Land Policy. \url{https://www.lincolninst.edu/publications/books/city-tech}
\end{thebibliography}

\end{document}
