\documentclass[10pt, twocolumn]{article}
\usepackage[margin=0.75in, top=0.6in, bottom=0.6in]{geometry}
\usepackage{newtxtext}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{xurl}
\usepackage{hyperref}
\usepackage{xcolor}

% Spacing
\setlist{nosep, leftmargin=*} 
\linespread{1.05} 
\setlength{\parskip}{3pt} 
\hyphenpenalty=10000
\exhyphenpenalty=10000
\renewcommand{\footnotesize}{\scriptsize}

% Section formatting with visual distinction
\titleformat{\section}{\normalsize\bfseries\scshape}{\thesection}{0.5em}{}[\vspace{1pt}\titlerule]
\titlespacing*{\section}{0pt}{6pt}{3pt}
\titleformat{\subsection}{\small\bfseries\itshape}{\thesubsection}{0.5em}{}
\titlespacing*{\subsection}{0pt}{4pt}{2pt}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    citecolor=black,
    breaklinks=true,
}

\pagestyle{plain} 

% Compact Bibliography
\makeatletter
\renewenvironment{thebibliography}[1]
     {\section*{References}%
      \scriptsize
      \setlength{\topsep}{0pt}
      \list{\@biblabel{\@arabic\c@enumiv}}%
           {\settowidth\labelwidth{\@biblabel{#1}}%
            \leftmargin\labelwidth
            \advance\leftmargin\labelsep
            \itemsep1pt\parsep0pt\parskip0pt
            \usecounter{enumiv}%
            \let\p@enumiv\@empty
            \renewcommand\theenumiv{\@arabic\c@enumiv}}%
      \sloppy
      \clubpenalty4000
      \@clubpenalty \clubpenalty
      \widowpenalty4000%
      \sfcode`\.\@m}
     {\def\@noitemerr
       {\@latex@warning{Empty `thebibliography' environment}}%
      \endlist}
\makeatother

\begin{document}

\twocolumn[{%
\noindent
\textbf{\Large Memo \#2: AI in City Services} \hfill \textbf{Daniel Hardesty Lewis} \\
\noindent
\textit{NYC's MyCity Chatbot: From Launch to Shutdown} \hfill March 27, 2026 \\
\textbf{PLAN A6613: AI and the Future of Cities} \hfill Spring 2026
\vspace{3pt}
\hrule
\vspace{8pt}
}]

\section*{I. Service Selection \& How It Works}

This memo examines New York City's \textbf{MyCity chatbot}, a generative AI assistant launched in October 2023 as the centerpiece of Mayor Eric Adams' AI Action Plan and described as ``the first of its kind for a major U.S. city'' (NYC Office of the Mayor, 2023). Powered by Microsoft Azure AI, the chatbot was designed to let business owners ``access trusted information from more than 2,000 NYC Business web pages'' covering ``compliance with codes and regulations, available business incentives, and best practices to avoid violations and fines'' (Lecher, 2024). The system used a retrieval-augmented generation architecture: a large language model, built on the same Microsoft infrastructure that underpins OpenAI's ChatGPT, synthesized responses from a corpus of NYC Department of Small Business Services web pages (Lecher, 2024). The interface presented these outputs as ``official NYC Business information,'' trained to provide authoritative guidance. Microsoft had previously partnered with Los Angeles to develop a similar business-assistance chatbot in 2017, though that earlier system's website is no longer available (Lecher, 2024). The MyCity chatbot built on the first phase of the portal, launched in March 2023 to help families check child care eligibility, explicitly extending the city's AI-driven approach into the far more consequential domain of regulatory compliance (NYC Office of the Mayor, 2023).

My firsthand testing of the chatbot, conducted before its February 2026 shutdown, exposed the system's fatal flaw. When asked straightforward navigational questions, such as how to register a business or find permit applications, the bot produced fluent, confident answers that linked to relevant city resources. The interface felt authoritative: city branding, official language, and a design engineered to simulate bureaucratic authority. But when asked nuanced regulatory questions involving tenant rights, labor law, or zoning compliance, the bot generated confident answers that could not be distinguished from accurate guidance. A small disclaimer noted the system ``may occasionally produce incorrect, harmful or biased content,'' yet the UI offered no way for users to separate accurate statutes from algorithmic fiction (Lecher, 2024). Users were encouraged to report inaccuracies through a feedback form that also asked them to rate their experience from one to five stars, a deeply cynical design choice that treated regulatory misinformation as a customer satisfaction issue (Lecher, 2024). The city stated on the chatbot's page that it would review submitted questions to improve answers and address ``harmful, illegal, or otherwise inappropriate'' content, but would otherwise delete user data within 30 days (Lecher, 2024). The central risk was structural rather than technical: the bot was most dangerous exactly when it appeared most helpful, weaponizing the city's authority to launder pure hallucinations.

Investigative testing by The Markup confirmed this observation at scale. The chatbot told business owners that employers could take workers' tips, violating New York Labor Law Section 196-d; that landlords could refuse tenants with Section 8 vouchers, despite source-of-income discrimination being illegal in New York City since 2008; and that cashless stores were permitted, contradicting a 2020 city law banning cashless businesses to prevent discrimination against unbanked customers (Lecher, 2024). The bot also claimed there were no regulations requiring businesses to inform employees about scheduling changes, contradicting the state Attorney General's published guidance on worker protections, and suggested it was acceptable for funeral homes to conceal service prices, violating the Federal Trade Commission's Funeral Rule (Lecher, 2024). When ten separate Markup reporters independently asked the same Section 8 question, the bot told all ten that landlords did not have to accept housing vouchers; at one point, it gave one reporter the correct answer, but this inconsistency itself illustrates the non-deterministic nature of the system (Lecher, 2024). These were not ambiguous interpretations but confident inversions of settled statute. Rosalind Black, Citywide Housing Director at Legal Services NYC, tested the bot herself after being alerted to The Markup's findings and found additional false information on housing, including that tenant lockouts were legal and that ``there are no restrictions on the amount of rent that you can charge a residential tenant,'' despite the extensive rent stabilization laws governing the city's housing stock (Lecher, 2024). Black warned that ``if this chatbot is not being done in a way that is responsible and accurate, it should be taken down'' (Lecher, 2024). Julia Stoyanovich, computer science professor and director of NYU's Center for Responsible AI, called the deployment ``reckless and irresponsible,'' arguing the city was ``rolling out software that is unproven without oversight'' (Offenhartz, 2024). Andrew Rigie, Executive Director of the NYC Hospitality Alliance, confirmed that a business owner had alerted him to inaccuracies, noting that while ``AI can be a powerful tool to support small business,'' it is ``a massive liability if it's providing the wrong legal information'' (Lecher, 2024).

\section*{II. Service Delivery: Benefits and Risks}

\subsection*{Effectiveness}

The Adams administration marketed the MyCity chatbot as a triumph of municipal innovation: 24/7 availability, instant responses, and multilingual accessibility, eliminating the need for business owners to navigate 2,000 pages of municipal code or wait for human support during business hours. By the time of its shutdown, the city's Office of Technology and Innovation claimed the bot had answered over 58,000 questions with a 95 percent accuracy rate (Vice, 2026). At face value, the tool vastly improved speed and accessibility over the status quo.

The 95 percent figure, however, obscures a critical problem of domain fitness. In most consumer applications, a 5 percent error rate is tolerable: a music recommendation engine that fails one in twenty times merely creates friction. In regulatory compliance, a 5 percent error rate is municipal malpractice. Five percent of 58,000 queries implies the city distributed approximately 2,900 instances of potentially harmful legal guidance to business owners who had no way to distinguish incorrect answers from correct ones. Internal audits and independent investigations continued to find persistent inconsistencies throughout 2024 and 2025, undermining the claimed accuracy rate (Vice, 2026). The comparison to a traditional approach is instructive: a human city employee might be slower and less available, but when that employee provides guidance on tenant rights or tip law, the answer is either correct or the employee knows to escalate. The chatbot manufactured the appearance of certainty without the substance of reliability.

\subsection*{Equity}

The chatbot's equity impact was entirely contradictory. On one hand, a free, multilingual, 24/7 tool reduces barriers for small business owners lacking the capital to hire attorneys for regulatory navigation, a distinct advantage in a city where roughly 3.2 million residents speak a non-English language at home. On the other hand, The Markup found that the exact same errors appeared when questions were asked in languages other than English (Lecher, 2024). The tool's multilingual capacity simply scaled its failures to non-English-speaking communities rather than insulating them. The populations most likely to rely on a free government bot over paid professional advice are precisely those least equipped to independently verify its legal outputs: immigrant entrepreneurs navigating unfamiliar regulatory systems, first-time business owners, and residents with limited familiarity with American labor law. The cashless-stores hallucination perfectly illustrates this trap: the 2020 law banning cashless businesses was enacted specifically to protect unbanked New Yorkers---disproportionately low-income residents and communities of color---from commercial exclusion. A chatbot that confidently tells a business owner this law does not exist actively dismantles a policy designed to protect the city's most vulnerable residents.

The UI worsened the trap by wrapping itself in ``official NYC Business information'' branding while appending a small disclaimer that responses should not be treated as professional or legal advice (Lecher, 2024; Fast Company, 2024). This structural deception required users to simultaneously trust the tool's authority and distrust its outputs, a cognitive burden demanding high legal literacy. Users with higher education and existing legal knowledge could parse this contradiction; those without could not. The result was a tool that offered the most help to those who needed it least and posed the greatest risk to those who needed it most.

\subsection*{City Workers}

The chatbot was inherently designed as an instrument of municipal austerity, sold under the politically palatable guise of labor augmentation. Standard public-sector AI frameworks pitch automation as a tool to handle rote work so human staff can focus on edge cases. MyCity inverted this dynamic. The system did not reduce municipal labor; it operated as a crude mechanism of cost exteriorization, displacing the unpaid burden of regulatory quality-assurance onto civil society. Rosalind Black at Legal Services NYC was forced to divert time from her primary, funded responsibilities to test and aggressively correct the bot's misinformation after being alerted by journalists (Lecher, 2024). Andrew Rigie, Executive Director of the NYC Hospitality Alliance, reported intervening after a business owner flagged the inaccuracies (Lecher, 2024). The city functionally conscripted community organizations and legal aid providers to serve as involuntary, unfunded quality-assurance teams for a tool that was supposed to render their advocacy unnecessary. Simultaneously, the city's own workers at OTI spent two years trapped in an unresolvable technical debt loop, attempting to salvage a system before its shutdown representing an ongoing public resource drain for a tool that never achieved fitness for purpose (Fast Company, 2024; Vice, 2026).

\section*{III. Viability \& Governance}

\subsection*{Financial Sustainability}

Taxpayers spent over \$600,000 to subsidize what was effectively an unconsented public beta test for Microsoft Azure AI over the chatbot's 28-month lifespan from October 2023 to February 2026 (Vice, 2026). Dividing this cost by the 58,000 questions the city claimed the bot answered yields a cost of approximately \$10.34 per query. This figure wildly underestimates the true cost. It excludes the salaries of OTI staff who spent two years attempting to patch a fundamentally flawed generative architecture, the institutional reputational damage to the city, and the unmeasured costs absorbed by civil society organizations forced to mitigate the bot's cascading misinformation. For comparison, New York City's 311 system handles roughly 25 million contacts per year. While 311's per-contact cost is difficult to isolate, the chatbot's per-query cost is vastly more expensive than the human-staffed alternative when quality-adjusted for the liability exposure of distributing thousands of potentially illegal answers. Mayor Mamdani cited the chatbot's sheer fiscal irresponsibility as a primary justification for its termination during a January 28, 2026 press conference addressing the city's \$12 billion budget gap, correctly diagnosing the tool as ``functionally unusable'' (Vice, 2026). By early February, the chatbot's webpage was quietly updated to read: ``The Chatbot beta test has ended.'' (Vice, 2026). Scaling this approach citywide would exponentially multiply both taxpayer burdens and liability surfaces, proving that municipal automation is not inherently cheaper than human expertise when the technology cannot meet the deterministic accuracy required by law.

\subsection*{Accountability}

The MyCity deployment was a masterclass in bureaucratic liability evasion. When the chatbot advised a business owner to violate Labor Law 196-d, the architecture ensured no individual or office could be held responsible. The liability chain fractured across multiple actors: Microsoft built the underlying Azure AI platform but declined to comment on its role (Lecher, 2024); OTI configured and deployed the system; the Department of Small Business Services owned the source corpus; and Mayor Adams championed the initiative for political capital. The entire system was designed to harness the state's authority while disavowing the state's liability. The Adams administration's defense was that the tool was a ``pilot program'' requiring real-world testing to ``iron out the kinks'' (Offenhartz, 2024; Fast Company, 2024). Leslie Brown, a spokesperson for OTI, stated that the city had been ``clear the chatbot is a pilot program'' that ``has already provided thousands of people with timely, accurate answers'' (Lecher, 2024). This framing forcibly conscripted citizens into the role of unconsenting beta testers of regulatory advice for 28 months. A Microsoft vice president, defending the company's own troubled Bing chatbot, had previously articulated the exact same extractive logic: ``You have to actually go out and start to test it with customers to find these kind of scenarios'' (Lecher, 2024). The city's AI Action Plan relied entirely on voluntary principles rather than enforceable standards (GovTech, 2025). Existing oversight mechanisms were functionally useless for this deployment: Local Law 144 (2021) required bias audits only for automated hiring tools, and Local Law 35 (2022) required only that city agencies report their use of automated decision systems, without mandating accuracy or safety thresholds (GovTech, 2025). Neither law would have prevented or detected the chatbot's catastrophic failure.

\subsection*{Looking Ahead}

The MyCity chatbot's lifecycle---from launch through controversy to shutdown---provoked a swift legislative backlash. In November 2025, the NYC City Council passed the Guaranteeing Unbiased AI Regulation and Disclosure Act, establishing an independent Office of Algorithmic Data Accountability with authority to audit and regulate AI tools used by city agencies, conduct pre-deployment assessments, investigate public complaints, and publish a public directory of all evaluated AI systems (GovTech, 2025). This legislation replaces voluntary principles with mandatory standards for fairness, privacy, transparency, and independent evaluation. At the state level, New York Senate Bill 7263 would create liability for chatbot operators, including government entities, that provide substantive advice equivalent to unauthorized practice of licensed professions, with a private right of action for damages (GovTech, 2025). If enacted, it would directly apply to future deployments.

If municipalities insist on deploying generative AI in compliance settings, three immediate guardrails are necessary to prevent similar liability disasters. First, \textbf{mandatory pre-deployment assessment} must apply to any AI system that delivers regulatory, legal, or compliance advice, requiring independent accuracy testing against a curated set of known statutory questions before public release. Second, \textbf{domain-specific accuracy thresholds} must distinguish between advisory AI---where a 95 percent accuracy rate may be acceptable---and compliance AI, where the threshold must approach 100 percent because every incorrect answer carries strict legal consequences for the user. Third, \textbf{structured escalation pathways} must ensure that every AI-generated response in a compliance context includes a clear, prominent mechanism to connect the user with a human expert, transforming the chatbot from an authoritative advisor into a triage tool that safely routes complex questions to qualified staff.\footnote{Full process documentation, including source downloads, line-by-line annotations, a cross-source synthesis matrix, and a claims verification audit, is available in the project repository at \texttt{memo2/}.}

\begin{thebibliography}{9}
\bibitem{fastco} Fast Company. (2024, April). NYC's AI chatbot is telling businesses to break the law. The city won't pull the plug. \textit{Fast Company}. \url{https://www.fastcompany.com/}
\bibitem{govtech} GovTech. (2025, November). New York City passes AI oversight legislation. \textit{Government Technology}. \url{https://www.govtech.com/artificial-intelligence/new-york-city-passes-ai-oversight-legislation}
\bibitem{markup} Lecher, C. (2024, March 29). NYC's AI Chatbot Tells Businesses to Break the Law. \textit{The Markup}. \url{https://themarkup.org/artificial-intelligence/2024/03/29/nycs-ai-chatbot-tells-businesses-to-break-the-law}
\bibitem{nyc2023} NYC Office of the Mayor. (2023, October 16). Mayor Adams Releases First-of-Its-Kind Plan for Responsible Artificial Intelligence Use in NYC Government. \url{https://www.nyc.gov/office-of-the-mayor/news/777-23/}
\bibitem{ap} Offenhartz, J. (2024, April 3). NYC's AI chatbot was caught telling businesses to break the law. The city isn't taking it down. \textit{AP News}. \url{https://apnews.com/article/new-york-city-chatbot-misinformation-6ebc71db5b770b9969c906a7ee4fae21}
\bibitem{vice} Vice. (2026, February). New York City kills AI chatbot after it kept telling people to break the law. \textit{Vice}. \url{https://www.vice.com/article/nyc-kills-ai-chatbot-mycity/}
\end{thebibliography}

\end{document}
